Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-06-20 manuscript.Rmd
responded to all requests

The first set of analyses use data from the proximal comparison to determine the "best case" accuracy of the models, training and testing on similar types of data. The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*. To assess the accuracy of body position classification for each recording session, we reserved the (temporally) last 25% of a session's proximal comparison data as the testing set. The testing set was never used as training data. The two types of models differed based on what data were used during training: A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75% of a recording session, then tested against the (temporally) last 25% of the same session. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification. 
<!-- data/reduce, duplicates of the group_split metrics but performed on reductions of sensor features from group_split_sensor_reduce.R -->

@@

Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded upright events?). High PPV means that we can be confident in the event label. Table \@ref(tab:metrics-by-class) shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types. Table \@ref(tab:metrics-table) shows the overall sensitivity and PPV across classes.

@@

In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period. A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to ~ 5 s in either direction). This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work [@YaoPlotz2019; @FranchakScott2021]. For each session, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding. Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.

Whereas the proximal analyses used all `r length(unique(ppts$id_uni))` sessions, this was not possible in the distal comparison. Because the start of the visit was scheduled during a time when the infant was awake, it was common for a nap to be needed following the proximal period. Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording. Three additional sessions were excluded because a caregiver accidentally turned off the video camera (*n* = 1) or (purposefully) left the house (*n* = 2). This left `r length(unique(agree$group$file))` sessions with usable distal comparison data.

@@

Visual inspection of Figure \@ref(fig:part2overall) indicated two extreme outliers, which we marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions. The "gray square" outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording. In contrast, the "gray diamond" outlier had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as "upright" but the models predicted it as "held". Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infants moved while walking upright.

What is notable about both outliers is that disagreements were restricted to a particular border case (supine vs. sitting; upright vs. held); accuracy for other classes remained strong. This suggests that their poor performance came as a result of spending a long time in an ambiguous position, not the result of the entire model failing to generalize to the later time period (or an error in sensor placement, such as if the parent removed the leggings and put them on backwards after a diaper change). To better capture the typical level of agreement, we report all correlations in Table \@ref(tab:pt2overalltable) excluding the two outliers. Overall agreement among the non-outlier sessions was excellent for both group models (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`). With the exception of held, within-class agreement was strong for the other four body positions.

@@

Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification. Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated the agreement analysis (correlating human-coded body position time with model-predicted time) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)). Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera. Bins were included only if there was > 7 minutes of usable data.

Table \@ref(tab:pt2binstable) shows the agreement correlation coefficients for group and individual models, including and excluding the two outliers identified in the previous section. Performance at a short timescale was similar to performance overall: Overall agreement after excluding outliers was excellent for both group (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`). Within-position correlations were weakest for held and strongest for upright regardless of the model type. Agreement for prone was better for group models, whereas sitting and supine were better predicted by individual models.

To describe the observed amount of prediction error in 10-minute bins, we subtracted the predicted duration (in minutes) for each body position in each bin from the human coded duration in that bin to create a prediction difference score. A score of 0 would indicate no error; positive differences indicate that the model overestimated the amount of time in a position, whereas negative differences indicate underestimation. Figure \@ref(fig:part2bins) plots the mean prediction difference for each session for each body position. The gray shaded area marks ± 1 minute of prediction error. The vast majority of session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation (with the previously-identified outliers as exceptions). For group models, we calculated the percentage of 10-min bins across participants that had errors < 1 minute: `r group_diff["Held","prop"]`% for held, `r group_diff["Supine","prop"]`% for supine, `r group_diff["Prone","prop"]`% for prone, `r group_diff["Sitting","prop"]`% for sitting, and `r group_diff["Upright","prop"]`% for upright. For individual models, the percent of 10-min bins with < 1 minute of error was: `r ind_diff["Held","prop"]`% for held, `r ind_diff["Supine","prop"]`% for supine, `r ind_diff["Prone","prop"]`% for prone, `r ind_diff["Sitting","prop"]`% for sitting, and `r ind_diff["Upright","prop"]`% for upright. Taken together, these results suggest that most estimates across models, sessions, and body positions had small amounts of prediction error even after a long delay separating the period that models were trained and tested.

@@

Figure \@ref(fig:timelines) depicts body position timelines across the day for each session, divided into younger (A) and older infants (B). Predictions from the group model were used to ensure that motion data were classified consistently across all sessions. Moreover, for infants who did not display all 5 body positions in the training period, group models were necessary to be able to predict those behaviors across the entire day. Session start times ranged from `r start_times[1]` to `r start_times[length(start_times)]` with a median of `r start_times[floor(length(start_times)/2)]`. Although the equipment was always dropped off in the morning, sometimes infants were unexpectedly asleep, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group). With two exceptions, recordings lasted until the infants' bedtime. Older participant \#1 had the equipment picked up on the same day rather than the next day, so the recording ended at 17:00. Older participant \#11 wore the equipment in the morning, but the family left the house at 10:15 and remained out for the rest of the day, choosing not to put the equipment back on when they returned home in the late afternoon. Among those participants who wore the equipment until bedtime, recordings ended between `r end_times[1]` and `r end_times[length(end_times)]` with a median of `r end_times[floor(length(end_times)/2)]`.

Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time. No caregiver mentioned needing to remove the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)). The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours. Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours. Body position data were available to describe `r apa_num(min(coverage$prop_accounted)*100, digits = 1)`-`r apa_num(max(coverage$prop_accounted)*100, digits = 1)`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted)*100, digits = 1)`%). Younger infant \#14 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home. Overall, the recordings produced a median of `r apa_num(median(coverage$hours_recorded))` hours of motion data, with the entire dataset totaling `r apa_num(sum(coverage$hours_recorded))` hours. Even the shortest full recording (`r apa_num(min(coverage$hours_recorded))` hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.

@@

In order to show the convergent validity of the classification method, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work [@Survey]. Table \@ref(tab:agetable) summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models. Past work using sparse EMA surveys across the day [@Survey] found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright. Prone time changed little. The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically. In @Survey, upright time was 0.6-5.5% of the time for infants 3-6 months, increasing to 22.0% at 12 months. Results from the current study were similar: younger infants' upright time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group. In @Survey, prone time was 2.9-9.2% of the time for infants 3-6 months, and was 7.2% at 12 months. In the current study, prone time was `r pos_sum_lookup["Prone", "Younger_Group"]`% for younger infants and `r pos_sum_lookup["Prone", "Older_Group"]`% for older infants. Held, sitting, and supine are more difficult to compare because prior work [@Survey] considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively. Further complicating the comparison, @Survey included a reclined category that we would classify as either supine or sitting (whichever it was closer to). Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \@ref(tab:agetable)). Sitting was greater for older infants (`r pos_sum_lookup["Sitting", "Older_Group"]`%) compared with younger infants (`r pos_sum_lookup["Sitting", "Younger_Group"]`%). Moreover, sitting was the most frequent body position followed by upright in 12-month-olds [@Survey], just as we observed in the current study. Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.

@@

Across three analyses, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors. Both group and individual models trained during the proximal comparison period performed well on testing data collected from the same recording period. Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes. Examining full recordings showed that the new method allows us to capture more data (median = `r apa_num(median(coverage$hours_recorded))` hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it. Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for capturing individual differences and describing developmental changes in motor behavior.

@@

How did classification in the current investigation compare to prior work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022]? Across body positions, median accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]*100`% for group models (Kappa = `r metrics_lookup["Kappa", "median_Group"]`) and `r metrics_lookup["Overall Accuracy", "median_Individual"]*100`% for individual models (Kappa = `r metrics_lookup["Kappa", "median_Individual"]`) during the proximal comparison. Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98%, kappas .93-.95), we note that those past values were obtained under ideal circumstances. A researcher applied the sensors in the lab [@FranchakScott2021; @AiraksinenRasanen2020] or home [@AiraksinenGallen2022] and set the stage for how infants and caregivers should interact. For example, both @FranchakScott2021 and @AiraksinenGallen2022 instructed the caregivers that the recorded sessions should mimic a "playtime", which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing). Examining the timelines in Figure \@ref(fig:timelines) shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing. With the exception of the brief guided session at the start of the recording (which ensured that every infant spent at least a short amount of time in each body position), the majority of the recorded behavior was truly natural and "in the wild".

The challenge only increased in the distal comparison period. Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV). Indeed, the distal comparison in Figure \@ref(fig:exemplar-timeline) contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch. With such a long delay in time from when the sensor garment was first put on, there is an increasing chance that sensors moved from their ideal positions. Regardless of these challenges, accuracy in the distal comparison continued to be strong. Barring two outliers (which we will discuss further), agreement was high for group (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`). Even at a finer timescale, most errors within 10-minute bins were less than 1 minute in total for all five body positions.

To put this in context, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders. Another comparison we can make is to the accuracy of the LENA&reg; device's automated metrics. Correlations between human and model frequencies were *r* = .698 for adult word count and *r* = .649 for child word count [@CristiaLavechin2020]. Even including outliers, correlations between human and model-predicted body position were *r* = `r group_agree_cor` (group) and *r* = `r split_agree_cor` (individual), stronger than the LENA&reg; automated counts. Note, however, that the LENA&reg; validation study used independently-collected data, whereas we are measuring performance for infants whose earlier data comprised the training set. More work is needed to know how well group models would predict a different set of infants' behavior in a different environmental and/or cultural context.

In their validation study, @CristiaLavechin2020 make the recommendation that researchers using LENA&reg; manually annotate a portion of data collected for their particular use case to ensure that automated LENA&reg; metrics are reliable for the participants and environment pertinent to the research question. Similarly, collecting some video data for each participant in future long-form motor behavior recordings provides a way to check reliability and identify errant predictions. Although predictions were accurate for most sessions, we noted two outliers for whom model predictions failed to match human annotations. Fortunately, we could reference video recordings to understand why the errors occurred. Idiosyncrasies in activities and device use make it challenging to decide *a priori* whether "odd" classifications (from the expectation of the researcher) are possible. For example, younger infant \#8 on Figure \@ref(fig:timelines) spent an unusual amount of time upright at an age where infants cannot yet stand. Inspecting the video revealed that the infant spent long periods of time suspended in a jumper that supported their body in an upright position. Although it is not feasible to collect large amounts of video data to check model predictions, interviewing caregivers about common activities and devices may provide a way to understand unexpected predictions.

Counterintuitively, the two outlier participants increased our confidence that the body position classification method works as intended. At first glance, seeing low agreement rates for two sessions would suggest that the models perform poorly at predicting some *infants'* behavior. Instead, we found that for those infants, errors were restricted to two positions during a single, long-lasting event, while the other three body positions continued to be classified correctly. In other words, the models failed to predict a particular *event* for each of the two infants. Unfortunately, those events lasted a long time leading to poor overall agreement. In the case of confusion between supine and sitting for the infant who was reclined, the confusion likely stems from the difficulty of categorizing behaviors that do not always fit neatly into discrete classes (e.g., continuous body position angles categorized as either supine or sitting).

@@

Aside from differences in validity, researchers may favor one modeling approach based on logistical concerns. Group models have several major advantages. First, by definition they are trained on more data, which might make predictions more consistent. Second, group models remove the need to get optimal training data from each participant. It can be difficult to elicit every behavior of interest in each infant; individual models require that every behavior that will be later predicted was displayed by each individual infant. In fact, group models remove the need to get *any* training data for a particular infant, meaning that they can be applied to any infant regardless of whether they have training data. Finally, applying a group model means that individual differences between infants are due to their behavior being classified in a consistent way, rather than having a separate set of rules for predicting each individual. However, given the variety in infant behavior, individual models are more agile in that they can capture the unique and unexpected. Likewise, individual models can be an excellent choice for rapid prototyping and pilot testing. A researcher can get proof-of-concept data from a single participant or a new kind of classification scheme (i.e., locomotion, restraint in device) without annotating an entire sample. Individual models may also be applied in clinical cases where infants vary widely in their motor abilities and/or use of assistive devices.

@@

In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn [@ClerkinHart2017; @KachergisYu2017; @RazAbney2019]. Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning. For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency [@ClerkinHart2017], but most objects are seen infrequently. Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music [@MendozaFausey2022; @CasillasElliott2021]. Long-form motor recordings provide a novel opportunity to measure the temporal structure in the sequences of body positions seen in Figure \@ref(fig:timelines). How skewed are infants' experiences with different motor behaviors, and how are they clustered in time? Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.

>>>>

The first set of analyses use data from the proximal comparison to determine the "best case" accuracy of the models, training and testing on similar types of data. The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*. To assess the accuracy of body position classification for each recording session, we reserved the (temporally) last 25% of a session's proximal comparison data as the testing set. The testing set was never used as training data, and was the same for both modeling approaches to facilitate direct comparisons between the models. We generated two different training datasets relative to each testing set. For the group model training set, we aggregated the first 75% of all **other** sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. The individual model training set used the first 75% of data from the testing set's session. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification. 
<!-- data/reduce, duplicates of the group_split metrics but performed on reductions of sensor features from group_split_sensor_reduce.R -->

@@

Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded upright events?). High PPV means that we can be confident in the event label. Table \@ref(tab:metrics-by-class) shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types.

@@

The first set of results showed that group and individual models trained from data during the proximal comparison period were within the proximal period. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period. A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. Distal videos had only coarse temporal synchrony with motion recordings which precluded calculating accuracy based on the proportion of matching events. Instead, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding [@YaoPlotz2019; @FranchakScott2021].

<!-- Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.  -->

Whereas the proximal analyses used all `r length(unique(ppts$id_uni))` sessions, this was not possible in the distal comparison. Because the start of the visit was scheduled during a time when the infant was awake, it was common for a nap to follow the proximal period. Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording. Three additional sessions were excluded because a caregiver accidentally turned off the video camera (*n* = 1) or (purposefully) left the house (*n* = 2). This left `r length(unique(agree$group$file))` sessions with usable distal comparison data.

@@

Visual inspection of Figure \@ref(fig:part2overall) indicated two extreme outliers, which we marked by a gray square and a gray diamond. The "gray square" outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position. This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording. In contrast, the "gray diamond" outlier had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as "upright" but the models predicted it as "held". Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infants moved while walking upright.

What is notable about both outliers is that disagreements were restricted to a particular border case (supine vs. sitting; upright vs. held); accuracy for other classes remained strong. This suggests that their poor performance came as a result of spending a long time in an ambiguous position, not the result of the entire model failing to generalize to the later time period (or an error in sensor placement, such as if the parent removed the leggings and put them on backwards after a diaper change). To better capture the typical level of agreement, we report all correlations in Table \@ref(tab:pt2overalltable) excluding the two outliers. Overall agreement among the non-outlier sessions was excellent for both group models (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`).

@@

Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated the agreement analysis after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)). Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera. Bins were included only if there was > 7 minutes of usable data. Table \@ref(tab:pt2binstable) shows the agreement correlation coefficients for group and individual models, including and excluding the two outliers identified in the previous section. Performance at a short timescale was similar to performance overall: Overall agreement after excluding outliers was excellent for both group (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`). Within-position correlations were weakest for held and strongest for upright regardless of the model type. Agreement for prone was better for group models, whereas sitting and supine were better predicted by individual models.

To describe the observed amount of prediction error in 10-minute bins, we subtracted the predicted duration (in minutes) for each body position in each bin from the human coded duration in that bin to create a prediction difference score. A score of 0 would indicate no error; positive differences indicate that the model overestimated the amount of time in a position, whereas negative differences indicate underestimation. Figure \@ref(fig:part2bins) plots the mean prediction difference for each session for each body position. The gray shaded area marks ± 1 minute of prediction error. Most session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation. For group models, we calculated the percentage of 10-min bins across participants that had errors < 1 minute: `r group_diff["Held","prop"]`% for held, `r group_diff["Supine","prop"]`% for supine, `r group_diff["Prone","prop"]`% for prone, `r group_diff["Sitting","prop"]`% for sitting, and `r group_diff["Upright","prop"]`% for upright. For individual models, the percent of 10-min bins with < 1 minute of error was: `r ind_diff["Held","prop"]`% for held, `r ind_diff["Supine","prop"]`% for supine, `r ind_diff["Prone","prop"]`% for prone, `r ind_diff["Sitting","prop"]`% for sitting, and `r ind_diff["Upright","prop"]`% for upright.

@@

Figure \@ref(fig:timelines) depicts body position timelines across the day for each session, divided into younger (A) and older infants (B). Predictions from the group model were used to ensure that motion data were classified consistently across all sessions. Moreover, for infants who did not display all 5 body positions in the training period, group models were necessary to predict those behaviors across the entire day. Session start times ranged from `r start_times[1]` to `r start_times[length(start_times)]` with a median of `r start_times[floor(length(start_times)/2)]`. Sometimes infants were unexpectedly asleep at the scheduled time, leading to a few sessions in which infants began wearing the sensors later than intended (such as \#11 in the younger group). With two exceptions, recordings lasted until the infants' bedtime. Older participant \#1 had the equipment picked up on the same day rather than the next day, so the recording ended at 17:00. Older participant \#11 wore the equipment in the morning, but the family left the house at 10:15 and remained out for the rest of the day, choosing not to put the equipment back on when they returned home in the late afternoon. Among those participants who wore the equipment until bedtime, recordings ended between `r end_times[1]` and `r end_times[length(end_times)]` with a median of `r end_times[floor(length(end_times)/2)]`.

Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time. No caregiver reporting removing the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)). The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours. Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours. Body position data were available to describe `r apa_num(min(coverage$prop_accounted)*100, digits = 1)`-`r apa_num(max(coverage$prop_accounted)*100, digits = 1)`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted)*100, digits = 1)`%). Younger infant \#14 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home. Overall, the recordings produced a median of `r apa_num(median(coverage$hours_recorded))` hours of motion data, with the entire dataset totaling `r apa_num(sum(coverage$hours_recorded))` hours. Even the shortest full recording (`r apa_num(min(coverage$hours_recorded))` hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.

@@

In order to show the convergent validity of the classification method, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work [@Survey]. Table \@ref(tab:agetable) summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models. EMA surveys [@Survey] found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright. The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically. In @Survey, upright time was 0.6-5.5% of the time for infants 3-6 months, increasing to 22.0% at 12 months. Results from the current study were similar: younger infants' upright time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group. In @Survey, prone time was 2.9-9.2% of the time for infants 3-6 months, and was 7.2% at 12 months. In the current study, prone time was `r pos_sum_lookup["Prone", "Younger_Group"]`% for younger infants and `r pos_sum_lookup["Prone", "Older_Group"]`% for older infants. Furthermore, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \@ref(tab:agetable)). Sitting was greater for older infants (`r pos_sum_lookup["Sitting", "Older_Group"]`%) compared with younger infants (`r pos_sum_lookup["Sitting", "Younger_Group"]`%). Moreover, sitting was the most frequent body position followed by upright in 12-month-olds [@Survey], just as we observed in the current study.

@@

Here, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors. Models trained during the proximal comparison period performed well on testing data collected from the same recording period. Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes. Examining full recordings showed that the new method allows us to capture more data (median = `r apa_num(median(coverage$hours_recorded))` hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it. Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.

@@

How did classification in the current investigation compare to prior work [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022]? Across body positions, median accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]*100`% for group models (Kappa = `r metrics_lookup["Kappa", "median_Group"]`) and `r metrics_lookup["Overall Accuracy", "median_Individual"]*100`% for individual models (Kappa = `r metrics_lookup["Kappa", "median_Individual"]`) during the proximal comparison. Although these accuracy and kappa values are slightly lower than in past work (accuracy 95-98%, kappas .93-.95), we note that those past values were obtained under ideal circumstances. A researcher applied the sensors in the lab [@FranchakScott2021; @AiraksinenRasanen2020] or home [@AiraksinenGallen2022] and set the stage for how infants and caregivers should interact. For example, both @FranchakScott2021 and @AiraksinenGallen2022 instructed the caregivers that the recorded sessions should mimic "playtime", which would encourage more common activities that would be easier to classify (e.g., crawling and sitting on the floor) and discourage idiosyncratic and potentially more challenging activities (e.g., positioning in infant-specific furniture like exersaucers and walkers, sitting or lying for long periods while eating or nursing). Examining the timelines in Figure \@ref(fig:timelines) shows how much variability there is between and within sessions in the earlier parts of the recording that were used for training and testing.

The challenge increased in the distal comparison period. Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV). Indeed, the distal comparison in Figure \@ref(fig:exemplar-timeline) contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch. Regardless of these challenges, accuracy in the distal comparison continued to be strong. Barring two outliers (which we will discuss further), agreement was high for group (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`). Even at a finer timescale, most errors within 10-minute bins were less than 1 minute in total for all five body positions. For comparison, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders.

Idiosyncrasies in activities and device use make it challenging to decide *a priori* whether "odd" classifications (from the expectation of the researcher) are possible. For example, younger infant \#8 on Figure \@ref(fig:timelines) spent an unusual amount of time upright at an age where infants cannot yet stand. Inspecting the video revealed that the infant spent long periods of time suspended in a jumper that supported their body in an upright position. Although it is not feasible to collect large amounts of video data to check model predictions, interviewing caregivers about common activities and devices may provide a way to understand unexpected predictions. Counterintuitively, the two outlier participants increased our confidence that the body position classification method works as intended. At first glance, seeing low agreement rates for two sessions would suggest that the models perform poorly at predicting some *infants'* behavior. Instead, we found that for those infants, errors were restricted to two positions during a single, long-lasting event, while the other three body positions continued to be classified correctly. In other words, the models failed to predict a particular *event* for each of the two infants.

@@

Aside from differences in validity, researchers may favor one modeling approach based on logistical concerns. Group models have several major advantages. First, by definition they are trained on more data, which might make predictions more consistent. Second, group models remove the need to get optimal training data from each participant. It can be difficult to elicit every behavior of interest in each infant; individual models require that every behavior that will be later predicted was displayed by each individual infant. In fact, group models remove the need to get *any* training data for a particular infant. Finally, applying a group model means that individual differences between infants are due to their behavior being classified in a consistent way, rather than having a separate set of rules for predicting each individual. However, given the variety in infant behavior, individual models are more agile in capturing the unique and unexpected. Likewise, individual models can be an excellent choice for rapid prototyping and pilot testing. A researcher can get proof-of-concept data from a single participant or a new kind of classification scheme (i.e., locomotion, restraint in device) without annotating an entire sample. Individual models may also be applied in clinical cases where infants vary widely in their motor abilities and/or use of assistive devices.

@@

In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn [@ClerkinHart2017; @KachergisYu2017; @RazAbney2019]. Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning. For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency [@ClerkinHart2017], but most objects are seen infrequently. Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music [@MendozaFausey2022; @CasillasElliott2021]. Long-form motor recordings provide a novel opportunity to measure the temporal structure in the sequences of body positions seen in Figure \@ref(fig:timelines). How skewed are infants' experiences with different motor behaviors, and how are they clustered in time? Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability. Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life.@MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices. Combining LENA&reg; and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.
